---
title: "Final Project"
date: “05/11/2020"
output:
  pdf_document: default
  html_document: default
  fig_width: 3
  fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

* Part1 USA national level analysis
* 1. Build Model 1 between USA cumulative positive cases and the date between 3/16/2020 to 3/29/2020.
```{r}
#1. Data pre-processing for training data of model 1 (M1)
library(data.table)
library(dplyr)
library(reshape2)
library(plyr)
library(ggplot2)
covid19_USA = fread('/Users/likehang/Desktop/lr final project/COVID-19_0408_USA.csv', data.table = FALSE) 
covid19_USA$date = as.Date(as.character(covid19_USA$date),tryFormats = "%Y%m%d")
covid19_USA = covid19_USA[order(as.Date(covid19_USA$date, format="%d/%m/%Y")),]
hist(covid19_USA$positive)
subset1 = covid19_USA$date>="2020-03-16" & covid19_USA$date <= "2020-03-29" 
train1_USA = covid19_USA[subset1,] %>% dplyr::select('date','positive')
summary(train1_USA)

#2. Build linear regression model 1 (M1)
M1 = lm(positive ~ date, data = train1_USA)
summary(M1)
anova(M1)
confint(M1)

#3. Assumption diagnosis of model 1 (M1)
##3.1 Prediction value vs residuals, test the assumption of linearity
predict = predict(M1, newdata = train1_USA)
plot(predict, M1$residuals)
##3.2 QQ Plot, test the assumption that error terms are normally distributed
qqnorm(M1$residuals)
##3.3 Residuals vs. Fits Plot, test the assumption of homoscedasticity
plot(M1$fitted.values, M1$residuals)
```

* 2. Build Model 2 between USA cumulative positive cases and the date between 3/16/2020 to 3/29/2020.
```{r}
# Because the M1 assumption tests are not satisfied, try transformed regression model 2 (M2)
#1. Transform "positive" variable and build transformed regression model 2
library(MASS)
boxcox_fit = boxcox(M1)
lam1 = boxcox_fit$x[which.max(boxcox_fit$y)]
lam1
train1_USA$positive = (train1_USA$positive^lam1-1)/lam1
M2 = lm(positive ~ date, data = train1_USA)
summary(M2)
anova(M2)
confint(M2)

#2. Assumption diagnosis
##2.1 Prediction value vs residuals, test the assumption of linearity
predict = predict(M2, newdata = train1_USA)
plot(predict, M2$residuals)
##2.2 QQ Plot, test the assumption that error terms are normally distributed
qqnorm(M2$residuals)
##2.3 Residuals vs. Fits Plot, test the assumption of homoscedasticity
plot(M2$fitted.values, M2$residuals)
```

* 3. Model 2 prediction for the date from 03/30/2020 to 04/08/2020
```{r}
#1. Create date sequence from 03/30/2020 to 04/08/2020 and make prediction by model 2
pred_date = seq.Date(from = as.Date('2020-03-30'), to = as.Date('2020-04-08'), by = 'days')
new_USA = data.frame(date = pred_date)
pred_USA = predict(M2, newdata = new_USA, interval = "prediction")
pred_USA

#2. Transform fitted values and prediction intervals to original scale
pred_USA = (pred_USA*lam1 + 1)^(1/lam1)
pred_USA = data.frame(pred_USA)

#3. Filter the test data for model 2 prediction and combind predict data, test data together
test_USA = covid19_USA[covid19_USA$date>="2020-03-30" & covid19_USA$date <= "2020-04-08", ] %>% dplyr::select('positive')
test_USA = cbind(new_USA, pred_USA, test_USA)

#4. Calculate prediction error
test_USA$pred_error = (test_USA$fit - test_USA$positive)^2
M2_pred_error = mean(test_USA$pred_error)
M2_pred_error

#5. Visualization of model 2 prediction
p <- ggplot(test_USA, aes(date, positive), xlab = "Date", ylab = "USA cumulative count of positive cases") +
  geom_point() 
p + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
    geom_line(aes(y = upr), color = "red", linetype = "dashed")+
    geom_line(aes(y = fit), color = "blue", linetype = "dashed")
```

* 4. Build Model 3 between USA cumulative positive cases and the date between 3/23/2020 to 3/29/2020.
```{r}
#1. Data processing for training data of model 3 (M4)
subset2 = covid19_USA$date>="2020-03-23" & covid19_USA$date <= "2020-03-29" 
train2_USA = covid19_USA[subset2,] %>% dplyr::select('date','positive')
summary(train2_USA)

#2. Build simple linear regression M3
M3 = lm(positive ~ date, data = train2_USA)
summary(M3)

#3. Transfrom variable "positive" and build transformed regression model 3 (M4)
library(MASS)
boxcox_fit = boxcox(M3)
lam2 = boxcox_fit$x[which.max(boxcox_fit$y)]
lam2
train2_USA$positive = (train2_USA$positive^lam2-1)/lam2
M4 = lm(positive ~ date, data = train2_USA)
summary(M4)
anova(M4)

#4. Assumption diagnosis for model 3 (M4)
##4.1 Prediction value vs residuals, test the assumption of linearity
predict = predict(M4, newdata = train2_USA)
plot(predict, M4$residuals)
##4.2 QQ Plot, test the assumption that error terms are normally distributed
qqnorm(M4$residuals)
##4.3 Residuals vs. Fits Plot, test the assumption of homoscedasticity
plot(M4$fitted.values, M4$residuals)
```

* 5. Model 3 prediction for the date from 03/30/2020 to 04/08/2020
```{r}
#1. Predict the cumulative count of positive cases between 3/30/2020 to 4/08/2020 by model 3 (M4)
pred2_USA = predict(M4, newdata = new_USA, interval = "prediction")
pred2_USA

#2. Transform fitted values and prediction intervals to original scale
pred2_USA = (pred2_USA*lam2 + 1)^(1/lam2)
pred2_USA = data.frame(pred2_USA)

#3. Filter the test data for model 3 (M4) prediction and combind predict data, test data together
test2_USA = covid19_USA[covid19_USA$date>="2020-03-30" & covid19_USA$date <= "2020-04-08", ] %>% dplyr::select('positive')
test2_USA = cbind(new_USA, pred2_USA, test2_USA)

#4. Calculate prediction error of model 3 (M4)
test2_USA$pred_error = (test2_USA$fit - test2_USA$positive)^2
M4_pred_error = mean(test2_USA$pred_error)
M4_pred_error

#5. Visualiztion of model 2 and model 3 prediction by ggplot2
names(test2_USA)[names(test2_USA)=="fit"] = "fit2"
names(test2_USA)[names(test2_USA)=="upr"] = "upr2"
names(test2_USA)[names(test2_USA)=="lwr"] = "lwr2"
test_USA = merge(test_USA, test2_USA, by = c("date","positive"))

p2 <- ggplot(data = test_USA, aes(x = date, y = positive)) + geom_point() +labs(x ="Date", y = "USA Cumulative COVID-19 positive cases")

p2 + geom_line(aes(y = lwr), color = "red", linetype = "dashed")+
    geom_line(aes(y = upr), color = "red", linetype = "dashed")+
    geom_point(aes(y = fit), color = "red", linetype = "dashed")+
    geom_line(aes(y = lwr2), color = "blue", linetype = "dashed")+
    geom_line(aes(y = upr2), color = "blue", linetype = "dashed")+
    geom_point(aes(y = fit2), color = "blue", linetype = "dashed")
```

* Part2 State level analysis
* 1. Select 5 states with top 5 cumulative positive cases on March 29th
```{r}
#1. Data pre-processing 
covid19_state = fread('/Users/likehang/Desktop/lr final project/COVID-19_0408_States.csv', data.table = FALSE) 
covid19_state$date = as.Date(as.character(covid19_state$date),tryFormats = "%Y%m%d")
covid19_state = covid19_state[order(as.Date(covid19_state$date,format="%d/%m/%Y")),]

#2. Find the top5 states with the largest cumulative count of positive cases on March 29, 2020
date_0329 = covid19_state[covid19_state$date == "2020-03-29",]
date_0329 = date_0329[order(date_0329[["positive"]], decreasing = TRUE)[1:5],]
date_0329$state
```

* 2. Build Model 4 between state cumulative positive cases and the date between 3/23/2020 to 3/29/2020.
```{r}
#1. #1. Data processing for training data of model 4 (M6)
subset1 = covid19_state$state %in% c('NJ', 'NY', "CA", "MI", "MA")
subset2 = covid19_state$date>="2020-03-23" & covid19_state$date <= "2020-03-29" 
train1_state= covid19_state[subset1 & subset2,] %>% dplyr::select('date','positive','state')
train1_state$state = as.factor(train1_state$state)

#2. Build simple linear regression M5
M5 = lm(positive ~ date + state, data = train1_state)
summary(M5)
anova(M5)

#3. Transfrom variable "positive" and build transformed regression model 4 (M6)
library(MASS)
boxcox_fit = boxcox(M5)
lam3 = boxcox_fit$x[which.max(boxcox_fit$y)]
lam3
train1_state$positive = (train1_state$positive^lam3-1)/lam3
M6 = lm(positive ~ date + state, data = train1_state)
summary(M6)
confint(M6)

#4. Assumption diagnosis for model 4 (M6)
##4.1 Prediction value vs residuals, test the assumption of linearity
predict = predict(M6, newdata = train1_state)
plot(predict, M6$residuals)
##4.2 QQ Plot, test the assumption that error terms are normally distributed
qqnorm(M6$residuals)
##4.3 Residuals vs. Fits Plot, test the assumption of homoscedasticity
plot(M6$fitted.values, M6$residuals)
```

* 5. Model 4 prediction for the date from 03/30/2020 to 04/08/2020 in each state
```{r}
#1. Predict the cumulative positive cases in each state between 3/30/2020 to 4/08/2020 by model 4 (M6)
pred_date = seq.Date(from = as.Date('2020-03-30'), to = as.Date('2020-04-08'), by = 'days')
state_5 = factor(rep(c('NJ', 'NY', "CA", "MI", "MA"), 10))
new_state = data.frame(date = rep(pred_date,each = 5), state = state_5)
pred1_state = predict(M6, newdata = new_state, interval = "prediction")
pred1_state

#2. Transform fitted values and prediction intervals to original scale
train1_state$positive = (train1_state$positive*lam3 + 1)^(1/lam3)
pred1_state = (pred1_state*lam3 + 1)^(1/lam3)
pred1_state = data.frame(pred1_state)

#3. Filter the test data for model 4 (M6) prediction and combind predict data, test data together
subset1 = covid19_state$state %in% c('NJ', 'NY', "CA", "MI", "MA")
subset2 = covid19_state$date>="2020-03-30" & covid19_state$date <= "2020-04-08" 
test1_state= covid19_state[subset1 & subset2,] %>% dplyr::select('positive','state', 'date')
test1_state = merge(cbind(new_state, pred1_state), test1_state, by=c("date","state"))
test1_state = test1_state[order(test1_state$state),]

test1_state$pred_error = (test1_state$fit - test1_state$positive)^2
M6_pred_error_CA = mean(test1_state$pred_error[test1_state$state == "CA"])
M6_pred_error_NY = mean(test1_state$pred_error[test1_state$state == "NY"])
M6_pred_error_NJ = mean(test1_state$pred_error[test1_state$state == "NJ"])
M6_pred_error_MI = mean(test1_state$pred_error[test1_state$state == "MI"])
M6_pred_error_MA = mean(test1_state$pred_error[test1_state$state == "MA"])
M6_pred_error_CA
M6_pred_error_NY
M6_pred_error_NJ
M6_pred_error_MI
M6_pred_error_MA

ratio_CA = sqrt(M6_pred_error_CA) / mean(test1_state$positive[test1_state$state == "CA"])
ratio_MI = sqrt(M6_pred_error_MI) / mean(test1_state$positive[test1_state$state == "MI"])
ratio_MA = sqrt(M6_pred_error_MA) / mean(test1_state$positive[test1_state$state == "MA"])
ratio_NY = sqrt(M6_pred_error_NY) / mean(test1_state$positive[test1_state$state == "NY"])
ratio_NJ = sqrt(M6_pred_error_NJ) / mean(test1_state$positive[test1_state$state == "NJ"])
ratio_CA
ratio_MA
ratio_MI
ratio_NY
ratio_NJ

test1_state = rbind.fill(test1_state, train1_state )
test1_state = test1_state[order(test1_state$state),]
```

```{r}
#4. Filter the test data for model 4 (M6) prediction and combind predict data, test data together
pred_date = seq.Date(from = as.Date('2020-03-23'), to = as.Date('2020-04-08'), by = 'days')
state_5 = factor(rep(c('NJ', 'NY', "CA", "MI", "MA"), 17))
new_state = data.frame(date = rep(pred_date,each = 5), state = state_5)
pred2_state = predict(M6, newdata = new_state, interval = "prediction")
pred2_state

subset1 = covid19_state$state %in% c('NJ', 'NY', "CA", "MI", "MA")
subset2 = covid19_state$date>="2020-03-23" & covid19_state$date <= "2020-04-08" 
test2_state= covid19_state[subset1 & subset2,] %>% dplyr::select('positive','state', 'date')
test2_state = merge(cbind(new_state, pred2_state), test2_state, by=c("date","state"))
test2_state = test2_state[order(test2_state$state),]

#5. Calculate prediction error of model 4 (M6) for each state
##5.1 MSPE
test2_state$pred_error = (test2_state$fit - test2_state$positive)^2
M6_pred_error_CA = mean(test2_state$pred_error[test2_state$state == "CA"])
M6_pred_error_NY = mean(test2_state$pred_error[test2_state$state == "NY"])
M6_pred_error_NJ = mean(test2_state$pred_error[test2_state$state == "NJ"])
M6_pred_error_MI = mean(test2_state$pred_error[test2_state$state == "MI"])
M6_pred_error_MA = mean(test2_state$pred_error[test2_state$state == "MA"])
M6_pred_error_CA
M6_pred_error_NY
M6_pred_error_NJ
M6_pred_error_MI
M6_pred_error_MA
#5.2 Ratio of root MSPE to mean cumulative positive cases in each state
ratio_CA = sqrt(M6_pred_error_CA) / mean(test2_state$positive[test2_state$state == "CA"])
ratio_MI = sqrt(M6_pred_error_MI) / mean(test2_state$positive[test2_state$state == "MI"])
ratio_MA = sqrt(M6_pred_error_MA) / mean(test2_state$positive[test2_state$state == "MA"])
ratio_NY = sqrt(M6_pred_error_NY) / mean(test2_state$positive[test2_state$state == "NY"])
ratio_NJ = sqrt(M6_pred_error_NJ) / mean(test2_state$positive[test2_state$state == "NJ"])
ratio_CA
ratio_MA
ratio_MI
ratio_NY
ratio_NJ

#6. Visualiztion of model 4 prediction by ggplot2
lp1 <- ggplot(data=test1_state, aes(x = date, y = positive, group= state, colour= state))+ geom_point() +labs(x ="Date", y = "Cumulative COVID-19 positive cases") + scale_y_continuous(breaks = seq(0, 300000, by = 5000)) + ylim(0, 300000) + geom_point(aes(y = fit), shape = 4) + geom_line(aes(y = lwr), linetype = "dashed") + geom_line(aes(y = upr), linetype = "dashed")
# Change the legend
lp1 + scale_shape_discrete(name  ="state",
                          breaks=c("NJ", "NY", "CA", "MI", "MA"),
                          labels=c("NJ", "NY", "CA", "MI", "MA"))

lp2 <- ggplot(data=test1_state, aes(x = date, y = positive, group= state, colour= state))+ geom_point() +labs(x ="Date", y = "Cumulative COVID-19 positive cases") + scale_y_continuous(breaks = seq(0, 300000, by = 5000)) + ylim(0, 300000) + geom_point(aes(y = fit), shape = 4) + geom_line(aes(y = lwr), linetype = "dashed") + geom_line(aes(y = upr), linetype = "dashed")
# Change the legend
lp2 + scale_shape_discrete(name  ="state",
                          breaks=c("NJ", "NY", "CA", "MI", "MA"),
                          labels=c("NJ", "NY", "CA", "MI", "MA"))
```

```{r}
##5.1 Visualiztion of model 4 prediction by plot
xmax = max(test1_state$date)
xmin = min(test1_state$date)
ymax = max(test1_state$fit)
ymin = min(test1_state$positive)

plot(test1_state$date, test1_state$positive, pch = 19, xlab = "Date", ylab = "Cumulative count of positive cases")
lines(test1_state$date, test1_state$positive, type = "p")
count = 1
for(current_state in c("NJ", "NY", "CA", "MI", "MA")){
  count = count + 1
  lines(test1_state$date[test1_state$state == current_state],
        test1_state$fit[test1_state$state == current_state], type = "p", col = count)
}
legend('bottomright', legend = c("NJ", "NY", "CA", "MI", "MA"), col = 1:5, lty = rep(1,5))
```


